---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tdcm

<!-- badges: start -->
[![R-CMD-check](https://github.com/cotterell/tdcm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/cotterell/tdcm/branch/main/graph/badge.svg)](https://app.codecov.io/gh/cotterell/tdcm?branch=main)
<!-- badges: end -->

The R package **tdcm** implements estimation of the transition DCM (TDCM; 
Madison & Bradshaw, 2018b), a longitudinal extension of the log-linear cognitive 
diagnosis model (LCDM; Henson, Templin, & Willse, 2009). As the LCDM subsumes 
many other DCMs, many other DCMs can be estimated longitudinally via the TDCM. 
The package includes functions to estimate the single-group and multigroup TDCM, 
summarize results of interest including item parameters, growth proportions, 
transition probabilities, transitional reliability, attribute correlations, 
model fit, and provide plots. 

## Installation

You can install the development version of `tdcm` like so:

``` {r, eval=FALSE}
# Install devtools from CRAN
install.packages("tdcm")
```

``` {r, eval=FALSE}
# Or the development version from GitHub:
# install.packages("devtools")
devtools::install_github("cotterell/tdcm")
```

## Getting Started

See `vignette("tdcm", package = "tdcm")` or for an overview of the `tdcm`
package.
