---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tdcm

<!-- badges: start -->
[![R-CMD-check](https://github.com/cotterell/tdcm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/cotterell/tdcm/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/cotterell/tdcm/branch/main/graph/badge.svg)](https://app.codecov.io/gh/cotterell/tdcm?branch=main)
<!-- badges: end -->

The '**tdcm**' R package lets users estimate the transition diagnostic 
classification model (TDCM) described in 
[Madison & Bradshaw (2018)](https://doi.org/10.1007/s11336-018-9638-5), a 
longitudinal extension of the log-linear cognitive diagnosis model (LDCM) in 
[Henson, Templin & Willse (2009)](https://doi.org/10.1007/s11336-008-9089-5). As 
the LCDM subsumes many other diagnostic classification models (DCMs), many other 
DCMs can be estimated longitudinally via the TDCM. The 'tdcm' package includes 
functions to estimate the single-group and multigroup TDCM, summarize results of 
interest including item parameters, growth proportions, transition 
probabilities, transitional reliability, attribute correlations, model fit, and 
growth plots.

## Installation

You can install the 'tdcm' package like so:

::: .pkgdown-release
``` {r, eval=FALSE}
# Install tdcm from CRAN
install.packages("tdcm")
```
:::

``` {r, eval=FALSE}
# Or install the latest development version from GitHub:
if (!require("devtools")) install.packages("devtools")
devtools::install_github("cotterell/tdcm")
```

## Getting Started

See `vignette("tdcm", package = "tdcm")` or for an overview of the 'tdcm'
package.
